### 2011-12-19 paper_draft.tex
added more citations

Genome Wide Association Studies (GWAS) are a comparatively easy and cheap way to find Single Nucleotide Polymorphisms (SNPs) which can be interesting because of their medical relevance. SNPs found through GWAS can be used to find candidate genes for a closer inspection or to predict disease risks. Genome Wide Association Studies make use of statistics to compare the alleles of patients to the alleles of healthy controls. By this the method does not allow to find causal differences but mere correlations. The first GWAS was published in 2005 and compared age-related macular degeneration in contrast to a healthy control group (doi:10.1126/science.1109557). Since the beginning the number of participants in those studies is rising and over 1200 GWAS have been performed (doi:10.1186/1471-2350-10-6.) and over 5000 SNPs have been linked to different diseases and traits in those studies %(http://www.genome.gov/page.cfm?pageid=26525384&clearquery=1#result_table).

@@

Although companies like 23andMe are willing to contribute to science it is not easy for individual scientists to access the data. This is mainly due to privacy concerns of the customers. Nevertheless there are individual customers who are willingly sharing their data. Most do so by uploading their data to their personal website or to open software repositories like GitHub. While this is makes it possible for scientists to access the data, it requires a lot of work to keep track of all new genotyping data that is available to the public. While projects like the SNPedia try to keep track of all the files, this still does not allow to perform GWAS, as the phenotypic information is not attached to the genetic information. Projects that attach the phenotype to the genetic information, like the Personal Genome Project, still don't allow for an easy re-use of the data, as they lack an advanced programming interface (API) or other methods by which researchers could download the data.

@@

Processes with a longer runtime, such as parsing the genotyping results, creating archives of results which are be mailed to users and queries to external resources are handled using the ruby-gem Resque and a Redis-server. Search-features on the platform itself are implemented using SOLR and the ruby-gem sunspot. Additionally data can be requested from openSNP using the Distributed Annotation System. The data for this is stored in a mySQL-database, the delivery of the data is done by using ProServer, which was modified by Gel et al. for use in easyDAS.

>>>>

Genome Wide Association Studies (GWAS) are a comparatively easy and cheap way to find Single Nucleotide Polymorphisms (SNPs) which can be interesting because of their medical relevance. SNPs found through GWAS can be used to find candidate genes for a closer inspection or to predict disease risks. Genome Wide Association Studies make use of statistics to compare the alleles of patients to the alleles of healthy controls. By this the method does not allow to find causal differences but mere correlations. The first GWAS was published in 2005 and compared age-related macular degeneration in contrast to a healthy control group \cite{Klein2005}. Since the beginning the number of participants in those studies is rising and over 1200 GWAS have been performed \cite{Johnson2009} and over 5000 SNPs have been linked to different diseases and traits in those studies %(http://www.genome.gov/page.cfm?pageid=26525384&clearquery=1#result_table).

@@

Although companies like 23andMe are willing to contribute to science it is not easy for individual scientists to access the data. This is mainly due to privacy concerns of the customers. Nevertheless there are individual customers who are willingly sharing their data. Most do so by uploading their data to their personal website or to open software repositories like GitHub. While this is makes it possible for scientists to access the data, it requires a lot of work to keep track of all new genotyping data that is available to the public. While projects like the SNPedia try to keep track of all the files \cite{Cariaso2011}, this still does not allow to perform GWAS, as the phenotypic information is not attached to the genetic information. Projects that attach the phenotype to the genetic information, like the Personal Genome Project, still don't allow for an easy re-use of the data, as they lack an advanced programming interface (API) or other methods by which researchers could download the data.

@@

Processes with a longer runtime, such as parsing the genotyping results, creating archives of results which are be mailed to users and queries to external resources are handled using the ruby-gem Resque and a Redis-server. Search-features on the platform itself are implemented using SOLR and the ruby-gem sunspot. Additionally data can be requested from openSNP using the Distributed Annotation System \cite{Dowell2001,Jenkinson2008}. The data for this is stored in a mySQL-database, the delivery of the data is done by using ProServer \cite{Finn2007}, which was modified by Gel et al. for use in easyDAS \cite{GelMoreno2011}.
